---
title: "Homework 1"
author: "Saryu Patel"
date: "9/22/2021"
output: github_document
---

```{r setup, message=FALSE}
library(tidyverse)
```

### Problem 1

```{r}
df <- tibble(
  rand_samp = rnorm(10),
  vec_logical = rand_samp > 0,
  vec_char = c("This", "is", "a", "character", "vector", "of", "length", "ten", "in", "df"),
  vec_factor = factor(c("green", "blue", "red", "green", "red", "green", "blue", "red", "green", "green"))
)
df
```

```{r}
mean(pull(df, 1))
mean(pull(df, 2))
mean(pull(df, 3))
mean(pull(df, 4))
```

Taking the mean of a numeric vector, such as the random sample from the standard Normal distribution, and the logical vector works. Taking the means of the character vector and the factor vector does not work.

```{r results='hide', warning=FALSE}
as.numeric(pull(df, 2))
as.numeric(pull(df, 3))
as.numeric(pull(df, 4))
```

The logical vector is converted into 1s and 0s because the vector is binary, so R assigns the value 1 to TRUE and 0 to FALSE. The character vector is converted into NAs because there is no way to assign a numeric value to characters. The factor vector is converted into 1s, 2s, and 3s because it is a factor with three levels, so R assigns a different value to each level. These results explain what happens when trying to take the mean. By treating the logical values as 1s and 0s, R is able to take the mean. Since there are no numerical values associated with the character vector, R is not able to take the mean. Even though there are numerical values associated with each of the factor levels, R cannot take the mean of the vector because the factor levels are treated as characters.

### Problem 2

```{r}
data("penguins", package = "palmerpenguins")
```

The column names are: `r colnames(penguins)`. 

Two important variables are species and island. The values in these columns are: `r unique(penguins$species)` and `r unique(penguins$island)` respectively.

The dataset spans the years `r unique(penguins$year)`.

The size of the dataset is `r nrow(penguins)` x `r ncol(penguins)`.

The mean flipper length is `r mean(penguins$flipper_length_mm, na.rm = TRUE)`.

```{r warning=FALSE}
ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm, color = species)) + 
  geom_point()
ggsave("bill_len_vs_flipper_len.pdf", height = 4, width = 4)
```

